---
title: "Construct design matrix"
author: "Jacqueline Gutman, Suvam Paul"
date: 'Last updated: `r format(Sys.Date(), "%B %d, %Y") ` '
output:
  html_document:
    toc: yes
    toc_depth: '4'
  html_notebook:
    highlight: tango
    theme: cosmo
    toc: yes
    toc_depth: 4
---

# Setup
## Load packages
## Read in database credentials and open connection

```{r setup, message=FALSE, echo = FALSE}
knitr::read_chunk("setup_notebook.R")
```

```{r setup_notebook}
```

```{r show_pkgs, echo = FALSE}
print_loaded_pkgs()
```

# Extract generated features from database

```{r}

"%ni%" = Negate( "%in%" )
in_school <- c("deidentified$generated$in_school_performance")
generated_features_admissions <- get_tbl_names_by_stage(edu_db_con, "deidentified", "generated") %>% 
  .[ . %ni% in_school ]

generated_features_admissions_inschool <- get_tbl_names_by_stage(edu_db_con, "deidentified", "generated")
    

```


```{r}
get_outcome_data <- function(outcome_name) {
  add_tbl_prefix(outcome_name, status = "outcome") %>% 
  put_tbl_to_memory(edu_db_con, .)
}


fetch_and_join <- function(df, tbl_name, conn, key) {
  tbl_name %>%
    put_tbl_to_memory(conn, .) %>%
    select(-one_of("appl_year")) %>%
    left_join(df, ., by = key)
}


create_outcome_features_data <- function (table_names, data) {
  reduce(table_names, fetch_and_join, 
       conn = edu_db_con, key = "study_id", 
       .init = data)
}
```

##AOA

```{r}
#get outcome
aoa <- get_outcome_data("aoa")

```


Admissions
```{r}
#create features and outcomes data
aoa_admissions <- create_outcome_features_data(generated_features_admissions, aoa)

#write to database
model_run_name <- "aoa_admissions"

aoa_admissions %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))
```



Admissions and in school

```{r}
#create features and outcome data
aoa_admissions_inschool <- create_outcome_features_data(generated_features_admissions_inschool, aoa)

#write to database
model_run_name <- "aoa_admissions_inschool"

aoa_admissions_inschool %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))
```

## STEP 1 


```{r, eval = FALSE}
#get outcomes
step1 <- get_outcome_data("step1_score")
```

Admissions
```{r, eval = FALSE}
#combine features with outcomes
step1_admissions <- create_outcome_features_data(generated_features_admissions, step1)

#write to database
model_run_name <- "step1_admissions"

step1_admissions %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))
```


Admissions and in school

```{r, eval = FALSE}
#combine features with outcomes
step1_admissions_inschool <- create_outcome_features_data(generated_features_admissions_inschool, step1)

#write to database
model_run_name <- "step1_admissions_inschool"

step1_admissions_inschool %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))

```



## STEP 2

```{r, eval = FALSE}
#get outcomes
step2 <- get_outcome_data("step2_score")
```

Admissions
```{r, eval = FALSE}
#combine features with outcomes
step2_admissions <- create_outcome_features_data(generated_features_admissions, step2)

#wrte to database
model_run_name <- "step2_admissions"

step2_admissions %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))

```


Admissions and in schools
```{r, eval = FALSE}

#combine features with outcomes
step2_admissions_inschool <- create_outcome_features_data(generated_features_admissions_inschool, step2)

#wrte to database
model_run_name <- "step2_admissions_inschool"

step2_admissions_inschool %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))

```

##Residency top 25

```{r}
#get outcomes
residency_top_25 <- get_outcome_data("residency_top25")
```


Admissions
```{r}

#combine features with outcomes
residency_top_25_admissions <- create_outcome_features_data(generated_features_admissions, residency_top_25)

#write to database
model_run_name <- "residency_top25_admissions"

residency_top_25_admissions %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))

```


Admissions and in schools
```{r}
#combine features with outcomes
residency_top_25_admissions_inschool <- create_outcome_features_data(generated_features_admissions_inschool, residency_top_25)

#wrte to database
model_run_name <- "residency_top25_admissions_inschool"

residency_top_25_admissions_inschool %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))
```


# Residency non top 25 and noncompete

```{r}
residency_noncompete <- get_outcome_data("residency_noncompete")
```

Admissions
```{r}

#combine features with outcomes
residency_noncompete_admissions <- create_outcome_features_data(generated_features_admissions,
                                                                residency_noncompete)

#write to database
model_run_name <- "residency_noncompete_admissions"

residency_noncompete_admissions %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))

```


Admissions and in schools
```{r}
#combine features with outcomes
residency_noncompete_admissions_inschool <- create_outcome_features_data(
  generated_features_admissions_inschool, residency_noncompete)

#write to database
model_run_name <- "residency_noncompete_admissions_inschool"

residency_noncompete_admissions_inschool %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))
```

# Worry score school outcomes

```{r}
worry_score <- get_outcome_data("worry_score_school_outcomes")
```

Admissions
```{r}

#combine features with outcomes
worry_score_admissions <- create_outcome_features_data(generated_features_admissions, worry_score)

#write to database
model_run_name <- "worry_score_admissions"

worry_score_admissions %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))

```


# Screeners

```{r}
screener_scores <- get_outcome_data("screened_apps")
```

Admissions
```{r}

#combine features with outcomes
screener_scores_admissions <- create_outcome_features_data(generated_features_admissions, screener_scores) %>% 
  select(-num_schools_accepted_to)

#write to database
model_run_name <- "screener_scores_admissions"

screener_scores_admissions %>%
  write_to_database_single(edu_db_con, add_tbl_prefix(model_run_name, status = "model_data"))

```


```{r}
disconnect_all()
```

